```{bash}
RETICULATE_MINICONDA_ENABLED=FALSE
```


```{python}
import os
os.sep
print(os.sep)
```

<!-- http://r-statistics.co/Linear-Regression.html -->
  
Assumptions of linear regression:
    1- Linearity: The relationship between X and the mean of Y is linear.
    2 -Homoscedasticity: The variance of residual is the same for any value of X.
    3- Independence: Observations are independent of each other.
Let's start with the built-in data set cars. cars data has 50 rows and two columns.

```{r}
setwd('/home/emre/fMRI/Codes_GH/DataScienceExamples')
head(cars)
str(cars)
```
# Exploratory Data Analysis

## Scatter Plot


```{r}
scatter.smooth(x=cars$speed, y=cars$dist, main="Dist ~ Speed",ylab = "Distance", xlab = "Speed")  # scatterplot
```
Plot shows there is a linear (the first assumption) and increasing relationship between two variables. 

## Box plot
```{r}
boxplot(cars$speed, main="Speed", sub=paste("Outlier rows: ", boxplot.stats(cars$speed)$out))  # box plot for 'speed'
boxplot(cars$dist, main="Distance", sub=paste("Outlier rows: ", boxplot.stats(cars$dist)$out))  # box plot for 'distance'
```



