---
title: "OLS Mathematical"
author: "H Emre Kale"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Now we will explore the mathematical derivation of ordinary least squares parameter estimation of the  linear models. If we want to express using matrix notation:

$$y = X\beta + \epsilon$$
$$\epsilon \sim N(0,\sigma)$$
where $y$ is the observation vector, $X$ is the design matrix, $\beta$'s are the coefficients and \epsilon is the error. Then we want to find $\beta$ such that it minimises the total error.

$$\arg\min_{\beta} \epsilon^T\epsilon \text{, where}  $$
$$ \epsilon = y - X\beta $$
and

$\beta estimates are $\beta$ values that minimize the mean squared error.

$$\hat{\boldsymbol{\beta}} = \underset{\boldsymbol{\beta}}{\operatorname{arg\,min}}\,(y - X\beta)^T(y - X\beta)$$
Let's write error in open form.

$$(y - X\beta)^T(y - X\beta) = y^Ty - y^TX\beta - (X\beta)^Ty+ (X\beta)^T X\beta$$

Remember $(AB)^T = B^TA^T$ so:

$$(y^T(X\beta))^T =  (X\beta)^T y$$
so the second and third term of the equation are transpose of each other and they are scalars. Since transpose of a scalar is the same scalar then, both of the terms are same so:

$$(y - X\beta)^T(y - X\beta) = y^Ty - 2(X\beta)^Ty + \beta^T X^T X\beta$$
This is a quadratic equation and in order to find the $\beta$ which minimizes this equation wrt $\beta$, we will take the partial derivative of the equation. We can find partial derivatives term by term.

We can easily see that the partial derivative is 0 for the first term:
$$\frac{\partial y^Ty}{\partial\beta} = 0$$
The open form of the second term is:

$$(X\beta)^T y = 
(\beta_1x_{11}y_1 +  \beta_2x_{12}y_1 + \cdots  \beta_px_{1p}y_1) +
(\beta_1x_{21}y_2 + \beta_2x_{22}y_2 + \cdots + \beta_px_{2p}y_2) + \cdots +
(\beta_1x_{N1}y_N + \beta_1x_{N2}y_N + \cdots + \beta_px_{Np}y_N)
$$

Then for eg. partial derivative of this equation wrt $\beta_1$ is:

$$\frac{\partial(X\beta)^T y}{\partial\beta_1} = x_{11}y_1 + x_{21}y_2 + \cdots + x_{N1}y_N$$ 

and in general form: 

$$\frac{\partial(X\beta)^T y}{\partial\beta_i} = x_{1i}y_1 + x_{2i}y_2 + \cdots + x_{Ni}y_N$$ 

then finally if put these terms in matrix form we see:

$$ 
\frac{\partial(X\beta)^T y}{\partial\beta} = \begin{bmatrix}
\frac{\partial(X\beta)^T y}{\partial\beta_1}\\
\frac{\partial(X\beta)^T y}{\partial\beta_2}\\
\vdots\\
\frac{\partial(X\beta)^T y}{\partial\beta_p} 
\end{bmatrix} 
= 
\begin{bmatrix}
x_{11}y_1 + x_{21}y_2 + \cdots + x_{N1}y_N \\x_{12}y_1 + x_{22}y_2 + \cdots + x_{N2}y_N \\  \vdots \\
x_{N2}y_1 + x_{N2}y_2 + \cdots + x_{Np}y_N 
\end{bmatrix}  = X^Ty
$$


For the third term let's define the vector m X 1 vector b $$\frac{\partial b^TAb}{\partial b}$$ and N x N matrix A. Where b is a one dimensional vector.

$$b = \begin{bmatrix}
b_1 & b_2 &\cdots& b_m \end{bmatrix}^T$$

and,

$$ A =
\begin{bmatrix}
a_{11} & a_{12} & \cdots & a_{1m} \\
a_{21} & a_{22} & \cdots & a_{2m} \\
\vdots & \vdots & \ddots & \vdots \\
a_{m1} & a_{m2} & \cdots & a_{mm}
\end{bmatrix} 
$$

and $a_{ij} = a_{ji}$ values are symmetric.

$$b^TAb = 
b_1(b_1a_{11} + b_2a_{12} + \cdots + b_Na_{1m}) + b_2(b_1a_{21} + b_2a_{22} + \cdots + b_ma_{2m}) + \cdots  \\
+b_m(b_1a_{m1} + b_2a_{m2} + \cdots + b_Na_{mm})
$$
As we did before partial derivative of this equation wrt $\beta_1$ is: 

$$\frac{\partial\beta^TA\beta}{\partial b_1} = 2b_1a_{11} + (b_2a_{12}+ b_3a_{13} + \cdots + b_ma_{1m}) + (b_2a_{21} + b_3a_{31} + \cdots + b_ma_{m1} )$$ 
Since $a_{ij} = a_{ji}$ values are symmetric :

$$\frac{\partial\beta^TA\beta}{\partial b_1} = 2b_1a_{11} + 2b_2a_{12}  +  \cdots + 2b_ma_{1m}$$

generally:
$$\frac{\partial\beta^TA\beta}{\partial b_i} = 2b_1a_{i1} + 2b_2a_{i2}  +  \cdots + 2b_ma_{im}$$
in matrix form: 

$$ 
\frac{\partial b^TAb}{\partial b} = \begin{bmatrix}
\frac{\partial b^TAb}{\partial b_1}\\
\frac{\partial b^TAb}{\partial b_2}\\
\vdots\\
\frac{\partial b^TAb)^T y}{\partial b_m} 
\end{bmatrix} = 
\begin{bmatrix}
2b_1a_{11} + 2b_2a_{12}  +  \cdots + 2b_ma_{1m}\\ 
2b_1a_{21} + 2b_2a_{22}  +  \cdots + 2b_ma_{2m}\\
\vdots \\
2b_1a_{m1} + 2b_2a_{m2}  +  \cdots + 2b_ma_{mm}\\
\end{bmatrix}  = 2Ab
$$
If $A = X^TX$ and $b = \beta$ into this equation we see $\frac{\partial \beta X^TX\beta}{\partial \beta} = 2 X^TX\beta$. Then equation becomes

$$X^Ty - X^TX\beta = 0 $$ and
$$ \beta = (X^TX)^{-1}X^Ty $$


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
